### 2024-05-29 Manuscript_Figures.Rmd
Consistent color scheme

---
title: "Manuscript Figures"
output: html_document
date: "2024-03-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(terra)
library(tidyterra)
library(ggthemes)
library(ggpattern)
library(sf)
```

## Visualize Burn Plot Pre-Landis

Find a way to neatly show initial communities. Show full landscape and burn plot. For full landscape, perhaps just color it by dominant tree species. Show more information for burn plot.

### Full Landis Run

Read in IC map and IC csv. Find tree species with highest biomass for each MapCode, then add that info to the raster.
```{r}
IC_map <- rast(here("1.LANDIS-MODEL","LANDIS_run-Fire2-Dry","IC_Map_FortBragg.tif"))
FB_shp <- vect(here("FtBraggBoundary","FtBraggBoundary.shp"))
FB_shp <- project(FB_shp, IC_map)
FB_shp <- crop(FB_shp, IC_map)

IC_file <- read.csv(here("1.LANDIS-MODEL","LANDIS_run-Fire2-Dry","community-input-file-0.csv"))
IC_dom <- IC_file %>%
  group_by(MapCode,SpeciesName) %>%
  summarize(total_sp_biomass = sum(CohortBiomass)) %>%
  ungroup() %>%
  group_by(MapCode) %>%
  slice_max(total_sp_biomass) %>%
  mutate(SpeciesName = case_when(SpeciesName=="LobPine" ~ "Loblolly Pine",
                                 SpeciesName=="LongleafPine" ~ "Longleaf Pine",
                                 SpeciesName=="SlashPine" ~ "Slash Pine",
                                 SpeciesName=="TulipTree" ~ "Tulip Poplar",
                                 SpeciesName=="RedMaple" ~ "Red Maple",
                                 SpeciesName=="SweetGum" ~ "Sweetgum",
                                 SpeciesName=="WhiteOak" ~ "Oak spp.",
                                 SpeciesName=="TurkeyOak" ~ "Oak spp.",
                                 SpeciesName=="ShortPine" ~ "Shortleaf Pine")) %>%
  mutate(SpeciesName = factor(SpeciesName,
                              levels = c(
                                "Oak spp.",
                                "Sweetgum",
                                "Red Maple",
                                "Slash Pine",
                                "Loblolly Pine",
                                "Longleaf Pine",
                                "Shortleaf Pine",
                                "Tulip Poplar"
                              ))) %>%
  mutate(SpeciesID = as.numeric(SpeciesName))
IC_map_df <- as.data.frame(IC_map) %>%
  rename(MapCode = IC_Map_FortBragg)
IC_map_dom <- left_join(IC_map_df, IC_dom) %>%
  mutate(SpeciesName = case_when(MapCode %in% c(0,1) ~ "Non-Forest",
                                 TRUE ~ SpeciesName)) %>%
  mutate(SpeciesID = if_else(SpeciesName=="Non-Forest",9,SpeciesID))

sp_names <- c(
  "Sweetgum",
  "Red Maple",
  "Oak spp.",
  "Slash Pine",
  "Loblolly Pine",
  "Longleaf Pine",
  "Shortleaf Pine",
  "Tulip Poplar",
  "Non-Forest"
)

IC_map_rst <- IC_map_dom %>% select(SpeciesID)
IC_map$IC_map_rst <- IC_map_rst
IC_dom_map <- IC_map$IC_map_rst
IC_dom_map <- IC_dom_map %>% mutate(IC_map_rst = factor(IC_map_rst, labels = sp_names))
IC_dom_map <- mask(IC_dom_map, FB_shp)
```

```{r}
map_colors <- c(colorblind_pal()(8)[c(2,3,1,4,5,6,7,8)],"gray70")
ggplot() +
  geom_spatraster(data=IC_dom_map) +
  scale_fill_manual(values=map_colors, na.value = NA) +
  labs(fill="Dominant Tree Species") +
  theme_bw()
ggsave("IC_map.jpg", path=here("Plots"), width = 10, height = 5)
```

### Zoom in on burn plot

Look at the dominant tree species
```{r}
bp4 <- vect(here("BurnPlots","FB_plot4.shp")) %>%
  project(crs(IC_dom_map))

ext_new = ext(bp4)
ext_new[1] = ext_new[1]-60
ext_new[2] = ext_new[2]+60
ext_new[3] = ext_new[3]-60
ext_new[4] = ext_new[4]+60
IC_crop_bp4 <- crop(IC_dom_map,ext_new)
```
```{r}
ggplot() +
  geom_spatraster(data=IC_crop_bp4)+
  scale_fill_colorblind()+
  geom_spatvector(data=bp4,fill=NA,linewidth=1)+
  theme_bw()
```

Look at forest types before and after

```{r}
before <- read.csv(here("1.LANDIS-MODEL","LANDIS_run-Fire2-Dry","community-input-file-0.csv"))
fire2_dry <- read.csv(here("1.LANDIS-MODEL","LANDIS_run-Fire2-Dry","community-input-file-50.csv"))
fire2_wet <- read.csv(here("1.LANDIS-MODEL","LANDIS_run-Fire2-Wet","community-input-file-50.csv"))
fire5_dry <- read.csv(here("1.LANDIS-MODEL","LANDIS_run-Fire5-Dry","community-input-file-50.csv"))
fire5_wet <- read.csv(here("1.LANDIS-MODEL","LANDIS_run-Fire2-Dry","community-input-file-50.csv"))
before$pre_post <- "Initial Conditions"
fire2_dry$pre_post <- "Fire2-Dry"
fire2_wet$pre_post <- "Fire2-Wet"
fire5_dry$pre_post <- "Fire5-Dry"
fire5_wet$pre_post <- "Fire5-Wet"
before_after <- rbind(before,fire2_dry,fire2_wet,fire5_dry,fire5_wet)
```

```{r}
BA_two_sp <- before_after %>%
  group_by(pre_post, MapCode, SpeciesName) %>%
  slice_max(CohortBiomass) %>%
  ungroup() %>%
  mutate(SpeciesName = case_when(SpeciesName=="LobPine" ~ "Loblolly Pine",
                                 SpeciesName=="LongleafPine" ~ "Longleaf Pine",
                                 SpeciesName=="SlashPine" ~ "Slash Pine",
                                 SpeciesName=="TulipTree" ~ "Tulip Poplar",
                                 SpeciesName=="RedMaple" ~ "Red Maple",
                                 SpeciesName=="SweetGum" ~ "Sweetgum",
                                 SpeciesName=="WhiteOak" ~ "White Oak",
                                 SpeciesName=="TurkeyOak" ~ "Turkey Oak",
                                 SpeciesName=="ShortPine" ~ "Shortleaf Pine")) %>%
  mutate(SpeciesName = factor(SpeciesName,
                              levels = c(
                                "Longleaf Pine",
                                "Loblolly Pine",
                                "Slash Pine",
                                "Shortleaf Pine",
                                "Tulip Poplar",
                                "Sweetgum",
                                "Red Maple",
                                "Oak spp."
                              ))) %>%
  group_by(pre_post,MapCode) %>%
  top_n(2, CohortBiomass) %>%
  arrange(MapCode, desc(CohortBiomass)) %>%
  mutate(rank = row_number()) %>%
  select(-CohortAge, -CohortBiomass) %>%
  pivot_wider(names_from = rank, values_from = SpeciesName) %>%
  rename(Highest_Biomass = `1`, Second_Highest_Biomass = `2`) %>%
  ungroup() %>%
  select(-`3`) %>%
  mutate(forest_type = if_else(
    is.na(Second_Highest_Biomass),
    Highest_Biomass,
    paste0("1. ",Highest_Biomass," 2. ",Second_Highest_Biomass)
    )) %>%
  select(MapCode,forest_type,pre_post) %>%
  mutate(forest_type = factor(forest_type))

forest_types <- unique(BA_two_sp$forest_type)
forest_types <- data.frame("forest_type" = forest_types)
forest_types$FT_ID = as.numeric(row.names(forest_types))
forest_types <- forest_types %>% mutate(forest_type = factor(forest_type))

BA_split <- BA_two_sp %>%
  pivot_wider(names_from = pre_post, values_from = forest_type) %>%
  left_join(forest_types, by=join_by("Initial Conditions"=="forest_type")) %>%
  rename(IC_ID = FT_ID) %>%
  left_join(forest_types, by=join_by("Fire2-Dry"=="forest_type")) %>%
  rename(F2D_ID = FT_ID) %>%
  left_join(forest_types, by=join_by("Fire2-Wet"=="forest_type")) %>%
  rename(F2W_ID = FT_ID) %>%
  left_join(forest_types, by=join_by("Fire5-Dry"=="forest_type")) %>%
  rename(F5D_ID = FT_ID) %>%
  left_join(forest_types, by=join_by("Fire5-Wet"=="forest_type")) %>%
  rename(F5W_ID = FT_ID) %>%
  mutate(IC_ID = as.numeric(IC_ID),
         F2D_ID = as.numeric(F2D_ID),
         F2W_ID = as.numeric(F2W_ID),
         F5D_ID = as.numeric(F5D_ID),
         F5W_ID = as.numeric(F5W_ID))

IC_map_df <- as.data.frame(IC_map) %>%
  rename(MapCode = IC_Map_FortBragg)
IC_df <- BA_split %>% select(MapCode,IC_ID)
F2D_df <- BA_split %>% select(MapCode,F2D_ID)
F2W_df <- BA_split %>% select(MapCode,F2W_ID)
F5D_df <- BA_split %>% select(MapCode,F5D_ID)
F5W_df <- BA_split %>% select(MapCode,F5W_ID)

Initial_map <- left_join(IC_map_df, IC_df)
F2D_map <- left_join(IC_map_df, F2D_df)
F2W_map <- left_join(IC_map_df, F2W_df)
F5D_map <- left_join(IC_map_df, F5D_df)
F5W_map <- left_join(IC_map_df, F5W_df)

IC_rst <- Initial_map %>% select(IC_ID)
F2D_rst <- F2D_map %>% select(F2D_ID)
F2W_rst <- F2W_map %>% select(F2W_ID)
F5D_rst <- F5D_map %>% select(F5D_ID)
F5W_rst <- F5W_map %>% select(F5W_ID)

IC_map$`Initial Conditions` <- IC_rst
IC_map$`Fire2-Dry` <- F2D_rst
IC_map$`Fire2-Wet` <- F2W_rst
IC_map$`Fire5-Dry` <- F5D_rst
IC_map$`Fire5-Wet` <- F5W_rst

IC_all <- IC_map$`Initial Conditions`
F2D_all <- IC_map$`Fire2-Dry`
F2W_all <- IC_map$`Fire2-Wet`
F5D_all <- IC_map$`Fire5-Dry`
F5W_all <- IC_map$`Fire5-Wet`

levels(IC_all) <- forest_types %>% select(FT_ID, forest_type)
levels(F2D_all) <- forest_types %>% select(FT_ID, forest_type)
levels(F2W_all) <- forest_types %>% select(FT_ID, forest_type)
levels(F5D_all) <- forest_types %>% select(FT_ID, forest_type)
levels(F5W_all) <- forest_types %>% select(FT_ID, forest_type)
before_after_all <- c(IC_all,F2D_all,F2W_all,F5D_all,F5W_all)
names(before_after_all) <- c("Initial Conditions","Fire2-Dry","Fire2-Wet","Fire5-Dry","Fire5-Wet")
```

```{r}
before_after_plot <- crop(before_after_all, ext_new)
my_colors = c(colorblind_pal()(8),"gray75")
ggplot() +
  geom_spatraster(data=before_after_plot) +
  facet_wrap(~lyr) +
  geom_spatvector(data=bp4, fill=NA, linewidth=1) +
  scale_fill_manual(values=my_colors, na.value = "transparent") +
  labs(fill="Most abundant tree\nspecies by biomass") +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())
```

### Stacked bar chart of species composition in burn plot
```{r}
bp_mapcodes <- unique(as.vector(values(crop(IC_map, bp4))))
BA_plot <- before_after %>% 
  filter(MapCode %in% bp_mapcodes) %>%
  group_by(pre_post, SpeciesName) %>%
  summarize(CohortAge_Avg = mean(CohortAge),
            CohortBiomass_Sum = sum(CohortBiomass)) %>%
  mutate(pre_post = factor(pre_post, levels = c("Initial Conditions",
                                                "Fire2-Dry",
                                                "Fire2-Wet",
                                                "Fire5-Dry",
                                                "Fire5-Wet"))) %>%
  mutate(SpeciesName = case_when(SpeciesName=="LobPine" ~ "Loblolly Pine",
                                 SpeciesName=="LongleafPine" ~ "Longleaf Pine",
                                 SpeciesName=="SlashPine" ~ "Slash Pine",
                                 SpeciesName=="TulipTree" ~ "Tulip Poplar",
                                 SpeciesName=="RedMaple" ~ "Red Maple",
                                 SpeciesName=="SweetGum" ~ "Sweetgum",
                                 SpeciesName=="WhiteOak" ~ "White Oak",
                                 SpeciesName=="TurkeyOak" ~ "Turkey Oak",
                                 SpeciesName=="ShortPine" ~ "Shortleaf Pine")) %>%
  mutate(SpeciesName = factor(SpeciesName,
                              levels = c(
                                "White Oak",
                                "Turkey Oak",
                                "Sweetgum",
                                "Red Maple",
                                "Tulip Poplar",
                                "Shortleaf Pine",
                                "Slash Pine",
                                "Loblolly Pine",
                                "Longleaf Pine"
                              )))

bar_colors <- colorblind_pal()(8)[c(2,3,1,4,5,6,7,8)]
stacked_barchart <- BA_plot %>%
  ggplot() +
  geom_bar(stat="identity",aes(pre_post,CohortBiomass_Sum, fill=SpeciesName)) +
  scale_fill_manual(values=bar_colors) +
  labs(x="LANDIS Run",
       y="Total Biomass (kg)",
       fill="Species") +
  theme_bw()
stacked_barchart

ggsave("biomass_barchart.jpg",path = here("Plots"), height=4)
```

### Plot changes in biomass and age distributions

```{r}
before_after %>%
  group_by(pre_post, MapCode, SpeciesName) %>%
  summarize(TotalBiomass = sum(CohortBiomass)) %>%
  ggplot() +
  geom_bar(aes(x=SpeciesName,y=TotalBiomass,fill=pre_post),stat = "identity", position = position_dodge()) +
  scale_fill_colorblind() +
  theme_bw()
```
```{r}
before_after %>%
  ggplot() +
  geom_boxplot(aes(x=SpeciesName,y=CohortAge,color=pre_post), outlier.shape = 1) + 
  scale_color_colorblind() +
  theme_bw()
```

## Visualize Treelists

```{r}
treelist <- read.csv(here("1.LANDIS-MODEL","LANDIS_run-Fire2-Dry","Treelist_postlandis_cycle0.csv"))
set.seed = 14
sample_mapcode <- sample(unique(treelist$MapCode),1)
mapcode <- treelist %>% filter(MapCode == sample_mapcode)
mapcodes <- treelist %>%
  filter((X>=100 & X<=300) & (Y>=500 & Y<=700))

mapcodes_vect <- vect(mapcodes, geom=c("X", "Y"), crs="", keepgeom=FALSE)

tree_polys <- list()
for(i in 1:nrow(mapcodes_vect)){
  tree_i <- mapcodes_vect[i,]
  buffer_i <- buffer(tree_i, width = mapcodes_vect$CD_m[i]/2)
  tree_polys[[i]] <- buffer_i
}
mapcodes_poly <- vect(tree_polys)
writeVector(mapcodes_poly, here("Plots","sample_mapcode_canopies.shp"))
mapcodes_poly <- mapcodes_poly %>%
  arrange(CD_m) %>%
  mutate(SpeciesName = case_when(SPECIES_SYMBOL=="PITA" ~ "Loblolly Pine",
                                 SPECIES_SYMBOL=="PIPA2" ~ "Longleaf Pine",
                                 SPECIES_SYMBOL=="PIEL" ~ "Slash Pine",
                                 SPECIES_SYMBOL=="LIST2" ~ "Tulip Poplar",
                                 SPECIES_SYMBOL=="ACRU" ~ "Red Maple",
                                 TRUE ~ "Other")) %>%
  mutate(SpeciesName = factor(SpeciesName,
                              levels = c(
                                "Other",
                                "Red Maple",
                                "Slash Pine",
                                "Loblolly Pine",
                                "Longleaf Pine",
                                "Tulip Poplar"
                              )))

new_colors <- colorblind_pal()(8)[c(1,4,5,6,8)]
example_mapcode <- ggplot() +
  geom_hline(yintercept = 600, linetype = "dashed") +
  geom_vline(xintercept = 200, linetype = "dashed") +
  geom_spatvector(data=mapcodes_poly, aes(fill=SpeciesName, color=SpeciesName), alpha=0.7) +
  scale_fill_manual(values = new_colors) +
  scale_color_manual(values = new_colors) +
  labs(x = "X (m)",
       y = "Y (m)",
       fill = "Species",
       color="Species") +
  theme_bw()
example_mapcode
ggsave("example_mapcode.jpg",path = here("Plots"), height=6)

#TODO: Replace species codes with species names
```